### 2021-07-31 supplement/supplement-main.Rmd
everything for supplement, including all analyses, figures, bibliography, word styles doc

---
title: "Beliefs about the development of mental life"
subtitle: "Supplemental materials"
author: "[authors removed to preserve anonymity]"
date: "July 31, 2021"
output:
  # word_document:
  #   reference_docx: word-styles-ref.docx
   bookdown::pdf_document2:
    latex_engine: xelatex
    fig_caption: yes
    toc: true
    toc_depth: 3
bibliography: "references.bib"
# cls: "article.cls"
csl: nature.csl
figsintext: yes
header-includes: 
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{float}
tables: yes
---

```{r setup, set.seed(94305), include = F}
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(include = F)
```

```{r resave plots}
# do you want to resave the plots?
# resave_plots <- T # uncomment if yes
resave_plots <- F # uncomment if no
```

```{r load packages}
library(ggdendro)
library(dendextend)
library(psych)
library(lme4)
library(mgcv)
library(kableExtra)
library(langcog)
library(parameters)
library(cowplot)
library(tidyverse)
```

```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", 
                                         chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", 
                                         chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", 
                                         chdir = TRUE)))
```

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four distinct suites of mental capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensations_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child’s life (Studies 2-3); and explored the lay theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

In these Supplemental Materials, we provide extended information about the results presented in the main paper, present additional results, and provide an item-by-item comparison of the capacity items used here to items used in previous work on mind perception [@gray.etal2007, @weisman.etal2017a, @weisman.etal2017b, @malle2019].

Note: This is a fully reproducible manuscript, written in RMarkdown using the "knitr" package for R [@xie2021]. The full, editable script for all analyses, figures, and tables (here and in the main text) is available in the .Rmd file that generated this document, which can be found in the following GitHub repository: [removed to preserve anonymity].

# Extended results

## Study 1

As described in the main text, we began with an expansive exploration of US adults’ attributions of mental life to infants and young children of different ages. We included a wide range of 60 capacities in this initial exploration, from basic physiological sensations, to capacities for perception, cognition, and emotion, to abilities for self-regulation and social interaction. We asked participants to assess these capacities at three different ages (at birth, 9 months, and 5 years), with the goal of characterizing the conceptual structure underlying participants’ reasoning about the development of mental life. Our primary analysis for this preregistered study was an exploratory factor analysis (EFA) of participants’ capacity ratings. This analysis allowed us to examine which capacities tend to "hang together" in participants’ assessments of the mental lives of infants and young children. Following previous work on mind perception, we argue that the suites of capacities revealed by EFA offer a meaningful approximation of the latent conceptual structure underlying participants’ reasoning about the developing human mind.

In the main text, we presented the four-factor EFA solution suggested by parallel analysis. Here, we present extended results for this EFA and the alternative EFA solutions suggested by minimizing BIC and by Weisman et al.'s factor retention criteria [@weisman.etal2017a], as well as a visualization of perceptions of development in these mental capacities.

### Exploratory factor analyses (EFAs)

In order to determine how many factors to retain, we examined the results of three factor retention protocols:

(1) Parallel analysis (presented in the main text), which compares the observed correlation structure to the correlation structure arising from random datasets of the same size; 
(2) Minimizing the Bayesian Information Criterion (BIC), which is one method of optimizing both goodness of fit and parsimony; and 
(3) A set of factor retention criteria that have been used in Weisman et al.’s previous work [@weisman.etal2017a], in which they retained factors with eigenvalues greater than 1.0, which individually accounted for greater than 5% of the shared variance before transformation or rotation, and which were the "dominant" factor (the factor with the strongest absolute factor loading) for at least one mental capacity after transformation.  (In this protocol, a factor had to meet all three of these criteria in order to be retained).

Our interpretation of how best to characterize our dataset (i.e., how many factors we observe) was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol. Here we present these three factor retention protocols separately.

All factor analyses were conducted using ordinary least squares to find the minimum residual solution via the "fa" function in the "psych" package for R [@revelle2021].

#### Parallel analysis

```{r s1 parallel analysis}
set.seed(94305)
s1_parallel <- fa.parallel(d1_all, plot = F)
```

```{r s1 efa parallel analysis}
s1_efa_par <- fa(d1_all, 
                 nfactors = s1_parallel$nfact, 
                 rotate = "oblimin")
# s1_efa_par %>% fa.sort()
```

As reported in the main text, parallel analysis suggested retaining `r s1_parallel$nfact` factors. We call these factors _cognition and control_, _social connection_, _bodily sensations_, and _negative affect_. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efaparload); see also Figure 1 (main text). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efaparvacc).

```{r tabs1efaparload, include = T, warning = F, message = F}
tabs1efaparload <- model_parameters(s1_efa_par, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F))) %>%
  rename(`Cog. Ctrl.` = MR1,
         `Soc. Conn.` = MR3,
         `Bod. Sens.` = MR2,
         `Neg. Afft.` = MR4)

tabs1efaparload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by parallel analysis: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+4]) %>%
  column_spec(3, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+5]) %>%
  column_spec(4, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+6]) %>%
  column_spec(5, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efaparvacc, include = T, warning = F, message = F}
s1_efa_par$Vaccounted %>% 
  data.frame() %>% 
  rename(`Cognition and control` = MR1,
         `Social connection` = MR3,
         `Bodily sensations` = MR2,
         `Negative affect` = MR4) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by parallel analysis: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r fig01 code, fig.width = 4.5, fig.asp = 1.2}
heatmap_fun(s1_efa_par, 
            factor_names = c("Cognition and control",
                             "Bodily sensations",
                             "Social connection",
                             "Negative affect"))

# save plot
if (resave_plots == T | !file.exists("../outputs/fig01.pdf")) {
  ggsave("../outputs/fig01.svg")
  ggsave("../outputs/fig01.pdf")
}
```

```{r fig01, fig.cap="Study 1, factor solution suggested by parallel analysis. (Identical to Figure 1 in main text.)"}
# knitr::include_graphics("../outputs/fig01.pdf")
```

#### Minimizing BIC

```{r s1 minimizing bic}
s1_minbic <- VSS(d1_all, plot = F)

s1_minbic <- s1_minbic$vss.stats %>%
  rownames_to_column("nfact") %>%
  mutate(nfact = as.numeric(nfact)) %>%
  filter(BIC == min(BIC)) %>%
  data.frame()
```

```{r s1 efa minimizing bic}
s1_efa_bic <- fa(d1_all, 
                 nfactors = s1_minbic$nfact, 
                 rotate = "oblimin") # does not converge even with max.iter >> 100000; max.iter does affect printout
# s1_efa_bic %>% fa.sort()
```

Minimizing BIC suggested retaining `r s1_minbic$nfact` factors. (Note, however, that this solution does not converge.) Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efabicload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efabicvacc).

```{r tabs1efabicload, include = T, warning = F, message = F}
tabs1efabicload <- model_parameters(s1_efa_bic, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F)))

tabs1efabicload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by minimizing BIC: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs1efabicload[,s1_minbic$nfact+4]) %>%
  column_spec(3, bold = tabs1efabicload[,s1_minbic$nfact+5]) %>%
  column_spec(4, bold = tabs1efabicload[,s1_minbic$nfact+6]) %>%
  column_spec(5, bold = tabs1efabicload[,s1_minbic$nfact+7]) %>%
  column_spec(6, bold = tabs1efabicload[,s1_minbic$nfact+8]) %>%
  column_spec(7, bold = tabs1efabicload[,s1_minbic$nfact+9]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efabicvacc, include = T, warning = F, message = F}
s1_efa_bic$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by minimizing BIC: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s1_efa_bic$loadings[]))[1]`") to resonate with the construct we have called _cognition and control_, the second ("`r names(data.frame(s1_efa_bic$loadings[]))[2]`") to resonate with the construct we have called _bodily sensations_, the third ("`r names(data.frame(s1_efa_bic$loadings[]))[3]`") to resonate with the construct we have called _social connection_, and the fourth ("`r names(data.frame(s1_efa_bic$loadings[]))[4]`") to resonate with the construct we have called _negative affect_. The fifth factor ("`r names(data.frame(s1_efa_bic$loadings[]))[5]`") appears to pick out capacities particularly relevant to social interaction (e.g., social learning, person recognition, humor, listening), and the sixth factor ("`r names(data.frame(s1_efa_bic$loadings[]))[6]`") appears to pick out negative emotions (e.g., anger, annoyance, boredom, social fear); however, there were no capacities that loaded strongly on either of these capacities.

```{r figS01 code, fig.width = 4.5, fig.asp = 1.2}
heatmap_fun(s1_efa_bic)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS01.pdf")) {
  ggsave("../outputs/figS01.svg")
  ggsave("../outputs/figS01.pdf")
}
```

```{r figS01, fig.cap="Study 1, factor solution suggested by minimizing BIC.", include = F}
knitr::include_graphics("../outputs/figS01.pdf")
```

#### Weisman et al.'s (2017) factor retention criteria

```{r s1 weisman et al criteria}
s1_weismanetal <- reten_fun(d1_all, rot_type = "oblimin")
s1_weismanetal
```

```{r s1 efa weisman et al criteria}
s1_efa_wdm <- fa(d1_all, 
                 nfactors = s1_weismanetal, 
                 rotate = "oblimin") 
# s1_efa_wdm %>% fa.sort()
```

Weisman et al.'s (2017) factor retention criteria suggested retaining `r s1_weismanetal` factors. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efawdmload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efawdmvacc).

```{r tabs1efawdmload, include = T, warning = F, message = F}
tabs1efawdmload <- model_parameters(s1_efa_wdm, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F)))

tabs1efawdmload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by Weisman et al.'s (2017) factor retention criteria: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs1efawdmload[,s1_weismanetal+4]) %>%
  column_spec(3, bold = tabs1efawdmload[,s1_weismanetal+5]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efawdmvacc, include = T, warning = F, message = F}
s1_efa_wdm$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by Weisman et al.'s (2017) factor retention criteria: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s1_efa_wdm$loadings[]))[1]`") to resonate most strongly with the construct we have called _cognition and control_, and the second ("`r names(data.frame(s1_efa_wdm$loadings[]))[2]`") to resonate most strongly with the construct we have called _bodily sensations_. These factors are also quite similar to the constructs of "agency" and "experience" in Gray et al.'s original studies of mind perception [@gray.etal2007].

```{r figS02 code, fig.width = 4.5, fig.asp = 1.2}
heatmap_fun(s1_efa_wdm)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS02.pdf")) {
  ggsave("../outputs/figS02.svg")
  ggsave("../outputs/figS02.pdf")
}
```

```{r figS02, fig.cap="Study 1, factor solution suggested by minimizing BIC."}
knitr::include_graphics("../outputs/figS02.pdf")
```

### Capacity ratings: Developmental trajectories

Study 1 also offered a first glimpse into US adults' perceptions of the development of mental life. Supplementary Figure \@ref(fig:figS03) provides a visualization of mean "scores" for each domain of mental life. To calculate these scores, we sorted each of the 60 mental capacities included in this study into on of the four domains (_bodily sensations_, _negative affect_, _social connection_ or _cognition and control_) according to the factor on which it loaded the most strongly and positively (see Supplementary Table \@ref(tab:tabcapcomp)); for each participant, we then took the average of their responses to capacities in each domain at each target age. As we go on to confirm in Studies 2-3, this visualization suggests that participants perceived capacities for _bodily sensations_ to be substantial at birth and to develop relatively little over the first five years of life; in contrast, they perceived capacities for _negative affect_, _social connection_, and in particular _cognition and control_ to be much more limited at birth and to development dramatically over the first five years of life.

```{r s1 scales}
d1_scales <- s1_efa_par$loadings[] %>%
  data.frame() %>%
  rownames_to_column("capacity") %>%
  gather(factor, loading, -capacity) %>%
  group_by(capacity) %>%
  top_n(1, abs(loading)) %>%
  ungroup() %>%
  mutate(factor = recode_factor(factor,
                                "MR2" = "Bodily sensations",
                                "MR4" = "Negative affect",
                                "MR3" = "Social connection",
                                "MR1" = "Cognition and control")) %>%
  arrange(factor, desc(loading)) %>%
  mutate(order = 1:nrow(.))
```

```{r s1 cap ratings dataset}
d1_cap_ratings <- d1_all %>%
  rownames_to_column("ResponseId_target") %>%
  separate(ResponseId_target, c("ResponseId", "target")) %>%
  mutate(target_year = recode(target,
                              target00mo = 0, # newborns
                              target0Xmo = 3/365, # 3 days
                              target01mo = 1/12, # 1 month
                              target02mo = 2/12, 
                              target04mo = 4/12, 
                              target06mo = 6/12, 
                              target09mo = 9/12,
                              target12mo = 12/12,
                              target18mo = 18/12,
                              target24mo = 24/12,
                              target36mo = 36/12,
                              target48mo = 48/12,
                              target60mo = 60/12)) %>%
    gather(capacity, response, -ResponseId, -starts_with("target")) %>%
  left_join(d1_scales %>% distinct(factor, capacity))
```

```{r s1 cap ratings summary stats}
d1_cap_ratings_sum <- d1_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup() %>%
  arrange(target_year, factor)

d1_cap_ratings_sum %>%
  mutate_if(is.numeric, funs(round_x(., 2)))
```

```{r s1 gam dataset}
d1_gam <- d1_cap_ratings %>%
  # implement correction for beta distribution
  mutate(response = case_when(response/100 < 0.005 ~ 0.005,
                              response/100 > 0.995 ~ 0.995,
                              TRUE ~ response/100))

contrasts(d1_gam$factor) <- cbind(bod_cog = c(1, 0, 0, 0),
                                  neg_cog = c(0, 1, 0, 0),
                                  soc_cog = c(0, 0, 1, 0))
```

```{r figS03 code, fig.width = 4, fig.asp = 0.5}
d1_gam %>%
  group_by(ResponseId, target_year, factor) %>%
  # # ALTERNATIVE: take just top 5 items per factor (like Study 2)
  # left_join(d1_scales %>% distinct(factor, capacity, loading)) %>%
  # top_n(5, loading) %>%
  summarise(score = mean(response, na.rm = T)) %>%
  ungroup() %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            # color = "#160072",
            alpha = 0.2, size = 0.02) +
  geom_path(data = . %>%
                    group_by(factor, target_year) %>%
                    summarise(mean = mean(score, na.rm = T)) %>%
                    ungroup,
            aes(y = mean),
            color = "black", lty = 2) +
            # color = "#160072",
  geom_pointrange(data = . %>%
                    group_by(factor, target_year) %>%
                    multi_boot_standard(col = "score") %>%
                    ungroup,
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black",
                  # color = "#160072",
                  fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response\n0 = not at all capable, 100 = completely capable") +
  guides(color = guide_none())

# save plot
if (resave_plots == T | !file.exists("../outputs/figS03.pdf")) {
  ggsave("../outputs/figS03.svg")
  ggsave("../outputs/figS03.pdf")
}
```

```{r figS03, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 1). Lighter lines represent individual participants’ scores, black points correspond to mean scores across the sample, and error bars are bootstrapped 95% confidence intervals.", include = T}
knitr::include_graphics("../outputs/figS03.pdf")
```

```{r s1 export results}
# save numerical results
s1_results <- list(demo = d1_demo,
                   efa = s1_efa_par,
                   efa_bic = s1_efa_bic,
                   efa_wdm = s1_efa_wdm,
                   cap_rat_ss = d1_cap_ratings_sum)

saveRDS(s1_results, "../outputs/s1_results.rds")
```

## Study 2

As described in the main text, in Study 2 we replicated the conceptual structure identified by Study 1 and then used it to chart how different aspects of mental life are perceived to change over development. The design of Study 2 was nearly identical to Study 1, except that instead of assessing 60 capacities for 3 target ages, each participant assessed 20 capacities for 13 target ages.

In the main text, we briefly described the four-factor EFA solution suggested by parallel analysis, and presented the results of a generalized additive model (GAM) predicting participants' responses as a function of target age, domain, and interactions between them. Here, we present the full results for the EFA mentioned in the main text, an alternative EFA solutions suggested by minimizing BIC, and the full results of the GAM.

### Exploratory factor analyses (EFAs)

As in Study 1, we examined the results of three factor retention protocols (parallel analysis, presented in the main text; minimizing BIC; and Weisman et al.’s factor retentia criteria [@weisman.etal2017a]), and our interpretation of how best to characterize our dataset was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol. Here we present these three factor retention protocols separately.

#### Parallel analysis

```{r s2 parallel analysis}
set.seed(94305)
s2_parallel <- fa.parallel(d2_all, plot = F)
```

```{r s2 efa parallel analysis}
s2_efa_par <- fa(d2_all, 
                 nfactors = s2_parallel$nfact, 
                 rotate = "oblimin")
# s2_efa_par %>% fa.sort()
```

As reported in the main text, parallel analysis suggested retaining `r s2_parallel$nfact` factors, and these factors were very similar to those found in Study 1 (_bodily sensations_, _negative affect_, _social connection_, and _cognition and control_); each item loaded most strongly on the factor it was selected to represent. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs2efaparload); see also Figure 1 (main text). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs2efaparvacc).

```{r tabs2efaparload, include = T, warning = F, message = F}
tabs2efaparload <- model_parameters(s2_efa_par, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F))) %>%
  rename(`Neg. Afft.` = MR1,
         `Bod. Sens.` = MR3,
         `Cog. Ctrl.` = MR2,
         `Soc. Conn.` = MR4)

tabs2efaparload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 2, factor solution suggested by parallel analysis: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+4]) %>%
  column_spec(3, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+5]) %>%
  column_spec(4, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+6]) %>%
  column_spec(5, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs2efaparvacc, include = T, warning = F, message = F}
s2_efa_par$Vaccounted %>% 
  data.frame() %>% 
  rename(`Negative affect` = MR1,
         `Bodily sensations` = MR3,
         `Cognition and control` = MR2,
         `Social connection` = MR4) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 2, factor solution suggested by parallel analysis: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r figS04 code, fig.width = 4.5, fig.asp = 0.5}
heatmap_fun(s2_efa_par, 
            factor_names = c("Negative affect",
                             "Cognition and control",
                             "Bodily sensations",
                             "Social connection"))

# save plot
if (resave_plots == T | !file.exists("../outputs/figS04.pdf")) {
  ggsave("../outputs/figS04.svg")
  ggsave("../outputs/figS04.pdf")
}
```

```{r figS04, fig.cap="Study 2, factor solution suggested by parallel analysis. (Identical to Figure 1 in main text.)"}
# knitr::include_graphics("../outputs/figS04.pdf")
```

#### Minimizing BIC

```{r s2 minimizing bic}
s2_minbic <- VSS(d2_all, plot = F)

s2_minbic <- s2_minbic$vss.stats %>%
  rownames_to_column("nfact") %>%
  mutate(nfact = as.numeric(nfact)) %>%
  filter(BIC == min(BIC)) %>%
  data.frame()
```

```{r s2 efa minimizing bic}
s2_efa_bic <- fa(d2_all, 
                 nfactors = s2_minbic$nfact, 
                 rotate = "oblimin") # does not converge even with max.iter >> 100000; max.iter does affect printout
# s2_efa_bic %>% fa.sort()
```

Minimizing BIC suggested retaining `r s2_minbic$nfact` factors. (Note, however, that this solution does not converge.) Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs2efabicload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs2efabicvacc).

```{r tabs2efabicload, include = T, warning = F, message = F}
tabs2efabicload <- model_parameters(s2_efa_bic, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F)))

tabs2efabicload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 2, factor solution suggested by minimizing BIC: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs2efabicload[,s2_minbic$nfact+4]) %>%
  column_spec(3, bold = tabs2efabicload[,s2_minbic$nfact+5]) %>%
  column_spec(4, bold = tabs2efabicload[,s2_minbic$nfact+6]) %>%
  column_spec(5, bold = tabs2efabicload[,s2_minbic$nfact+7]) %>%
  column_spec(6, bold = tabs2efabicload[,s2_minbic$nfact+8]) %>%
  column_spec(7, bold = tabs2efabicload[,s2_minbic$nfact+9]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs2efabicvacc, include = T, warning = F, message = F}
s2_efa_bic$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 2, factor solution suggested by minimizing BIC: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s2_efa_bic$loadings[]))[1]`") to resonate with the construct we have called _cognition and control_, the second ("`r names(data.frame(s2_efa_bic$loadings[]))[2]`") to resonate with the construct we have called _bodily sensations_, the third ("`r names(data.frame(s2_efa_bic$loadings[]))[3]`") to resonate with the construct we have called _social connection_, and the fourth and fifth ("`r names(data.frame(s2_efa_bic$loadings[]))[4]`" and "`r names(data.frame(s2_efa_bic$loadings[]))[5]`") to resonate with the construct we have called _negative affect_. The sixth factor ("`r names(data.frame(s2_efa_bic$loadings[]))[6]`") appears to pick out capacities particularly relevant to social interaction (e.g., social learning, love). We have no strong intuitions about how to interpret the seventh and eighth factors ("`r names(data.frame(s2_efa_bic$loadings[]))[7]`" and "`r names(data.frame(s2_efa_bic$loadings[]))[8]`").

```{r figS05 code, fig.width = 4.5, fig.asp = 1.2}
heatmap_fun(s2_efa_bic)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS05.pdf")) {
  ggsave("../outputs/figS05.svg")
  ggsave("../outputs/figS05.pdf")
}
```

```{r figS05, fig.cap="Study 2, factor solution suggested by minimizing BIC.", include = F}
knitr::include_graphics("../outputs/figS05.pdf")
```

#### Weisman et al.'s (2017) factor retention criteria

```{r s2 weisman et al criteria}
s2_weismanetal <- reten_fun(d2_all, rot_type = "oblimin")
s2_weismanetal
```

```{r s2 efa weisman et al criteria}
s2_efa_wdm <- fa(d2_all, 
                 nfactors = s2_weismanetal, 
                 rotate = "oblimin") 
# s2_efa_wdm %>% fa.sort()
```

Weisman et al.'s (2017) factor retention criteria suggested retaining `r s2_weismanetal` factors—i.e., the same `r s2_weismanetal`-factor solution suggested by parallel analysis. (See Supplementary Tables \@ref(tab:tabs2efaparload) and \@ref(tab:tabs2efaparvacc).)

### Capacity ratings: Developmental trajectories

Our primary goal in Study 2 was to chart perceptions of the development of mental life over the first five years. As described in the main text, we preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 2:

(1) We predicted that, on average, participants’ perception of children’s mental capacities would increase as a function of the target child’s age, with participants generally attributing more and greater abilities to older children.
(2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensations_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes).

To test these hypotheses, we conducted a multilevel beta regression via a generalized additive model, regressing item-level responses onto the domain (i.e., which of the four "factors" this capacity was selected to represented) and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: gamm(response ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1). Attempting to model random slopes for domain or for target age resulted in model convergence problems. Responses were rescaled to range from 0.005 to 0.995 in order to satisfy the constraints of the beta distribution. Domain was dummy-coded with _cognition and control_ as the baseline. The full results of this model are presented in Supplementary Tables \@ref(tab:tabs2gamfixed) and \@ref(tab:tabs2gamsmooth); see Figure 2, Panel A (main text) for a visualization of these perceived trajectories.

In brief, we view these results as confirming our hypotheses that US adults would rate older children as more capable across the board, but that perceptions of the development of children’s capacities would differ across domains. (See main text for extended interpretation.)

```{r s2 scales}
d2_scales <- s2_efa_par$loadings[] %>%
  data.frame() %>%
  rownames_to_column("capacity") %>%
  gather(factor, loading, -capacity) %>%
  group_by(factor) %>%
  top_n(5, loading) %>%
  ungroup() %>%
  mutate(factor = recode_factor(factor,
                                "MR3" = "Bodily sensations",
                                "MR1" = "Negative affect",
                                "MR4" = "Social connection",
                                "MR2" = "Cognition and control")) %>%
  arrange(factor, desc(loading)) %>%
  mutate(order = 1:nrow(.))
```

```{r s2 cap ratings dataset}
d2_cap_ratings <- d2_all %>%
  rownames_to_column("ResponseId_target") %>%
  separate(ResponseId_target, c("ResponseId", "target")) %>%
  mutate(target_year = recode(target,
                              target00mo = 0, # newborns
                              target0Xmo = 3/365, # 3 days
                              target01mo = 1/12, # 1 month
                              target02mo = 2/12, 
                              target04mo = 4/12, 
                              target06mo = 6/12, 
                              target09mo = 9/12,
                              target12mo = 12/12,
                              target18mo = 18/12,
                              target24mo = 24/12,
                              target36mo = 36/12,
                              target48mo = 48/12,
                              target60mo = 60/12)) %>%
    gather(capacity, response, -ResponseId, -starts_with("target")) %>%
  left_join(d2_scales %>% distinct(factor, capacity))
```

```{r s2 cap ratings summary stats}
d2_cap_ratings_sum <- d2_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup() %>%
  arrange(target_year, factor)

d2_cap_ratings_sum %>%
  mutate_if(is.numeric, funs(round_x(., 2)))
```

```{r s2 gam dataset}
d2_gam <- d2_cap_ratings %>%
  # implement correction for beta distribution
  mutate(response = case_when(response/100 < 0.005 ~ 0.005,
                              response/100 > 0.995 ~ 0.995,
                              TRUE ~ response/100))

contrasts(d2_gam$factor) <- cbind(bod_cog = c(1, 0, 0, 0),
                                  neg_cog = c(0, 1, 0, 0),
                                  soc_cog = c(0, 0, 1, 0))
```

```{r s2 gam}
if (!file.exists("../outputs/s2_gam.RDS")) {
  # takes ~20 minutes to run
  s2_gam <- gamm(response ~ s(target_year, by = factor) + factor, 
                 random = list(ResponseId = ~1,
                               capacity = ~1),
                 data = d2_gam,
                 family = betar,
                 method = "REML",
                 niterPQL = 40)

  saveRDS(s2_gam, "../outputs/s2_gam.RDS")

} else {
  s2_gam <- readRDS("../outputs/s2_gam.RDS")
}

summary(s2_gam$gam)
# model_parameters(s2_gam$gam)
```

```{r tabs2gam}
tabs2gam <- model_parameters(s2_gam$gam)
```

```{r tabs2gamfixed, include = T}
tabs2gam %>% 
  filter(Component == "conditional") %>%
  select(-c(df, df_error, Component)) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Cognition and control (intercept)",
    grepl("bod", Parameter) ~ "Bodily sensations vs. Cognition and control",
    grepl("neg", Parameter) ~ "Negative affect vs. Cognition and control",
    grepl("soc", Parameter) ~ "Social connection vs. Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "Study 2, GAM results: Parametric coefficients (fixed effects).",
        digits = 2, align = c("lrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r tabs2gamsmooth, include = T}
tabs2gam %>% 
  filter(Component == "smooth_terms") %>%
  select(-c(Coefficient, SE, starts_with("CI"), df_error, Component)) %>%
  mutate(Parameter = case_when(
    grepl("Bod", Parameter) ~ "Target year: Bodily sensations",
    grepl("Neg", Parameter) ~ "Target year: Negative affect",
    grepl("Soc", Parameter) ~ "Target year: Social connection",
    grepl("Cog", Parameter) ~ "Target year: Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  rename(`Estimated df` = df) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "Study 2, GAM results: Approximate significance of smooth terms.",
        digits = 2, align = c("lrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r s2 gam predictions}
d2_gam_r1_pred <- d2_gam %>%
  mutate(pred_score = predict.gam(s2_gam$gam, type = "response", 
                                  newdata = d2_gam)) %>%
  group_by(ResponseId, factor, target_year) %>%
  summarise(score = mean(response, na.rm = T),
            pred_score = mean(score, na.rm = T)) %>%
  ungroup()
```

```{r fig02a, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 2). Lighter lines represent individual participants’ scores, black points correspond to mean scores across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive model (beta regression).", fig.width = 4, fig.asp = 0.5}
d2_gam_r1_rsq <- round_x(summary(s2_gam$gam)$r.sq, 2)

fig02a <- d2_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            # color = "#160072",
            alpha = 0.2, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(data = . %>%
                    group_by(factor, target_year) %>%
                    multi_boot_standard(col = "score") %>%
                    ungroup,
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black",
                  # color = "#160072",
                  fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response\n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d2_gam_r1_rsq)))) +
  guides(color = guide_none())

# # save plot
# if (resave_plots == T | !file.exists("../outputs/fig02a.pdf")) {
#   ggsave("../outputs/fig02a.svg")
#   ggsave("../outputs/fig02a.pdf")
# }
```

```{r s2 export results}
# save numerical results
s2_results <- list(demo = d2_demo,
                   efa = s2_efa_par,
                   efa_bic = s2_efa_bic,
                   efa_wdm = s2_efa_wdm,
                   cap_rat_ss = d2_cap_ratings_sum,
                   gam = s2_gam)

saveRDS(s2_results, "../outputs/s2_results.rds")
```

## Study 3

As described in the main text, Study 3 was designed with two goals in mind: First, we aimed to provide a more direct assessment of the differences in perceived developmental trajectories surfaced by Study 2. Second, we aimed to probe the lay theories that might underlie the differentiation of the four factors surfaced by Studies 1-2 and the perceived developmental trajectories surfaced by Study 2 (and the current study).

In the main text, we presented the results of a generalized additive model (GAM) predicting participants' capacity ratings as a function of target age, domain, and interactions between them; briefly described the dimensionality reduction analyses we used to group developmental mechanisms into two types ("intrinsic" and "extrinsic"); presented the results of a multilevel linear regression model predicting participants' ratings of developmental mechanisms as a function of mechanism type, domain, and an interaction between them; briefly mentioned additional models of each domain separately; and provided visualizations of participants' selections of the "most important" mechanism for each capacity, by domain. Here, we present the full results of all of these models, as well as the results of a multilevel logistic regression model of the selection of "most important" mechanism.

### Capacity ratings: Developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 3 (identical to our hypotheses for Study 2):

(1) We predicted that, on average, participants’ perception of children’s mental capacities would increase as a function of the target child’s age, with participants generally attributing more and greater abilities to older children.
(2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensations_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes).

Following Study 2, to test these hypotheses we conducted a multilevel beta regression via a generalized additive model, regressing item-level responses onto the domain and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: gamm(responses ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1). Again, attempting to model random slopes for domain or for target age resulted in model convergence problems. Again, responses were rescaled to range from 0.005 to 0.995, and domain was dummy-coded with _cognition and control_ as the baseline. The full results of this model are presented in Supplementary Tables \@ref(tab:tabs3gamfixed) and \@ref(tab:tabs3gamsmooth); see Figure 2, Panel B (main text) for a visualization of these perceived trajectories.

In brief, we view these results as confirming our hypotheses that US adults would rate older children as more capable across the board, but that perceptions of the development of children’s capacities would differ across domains. (See main text for extended interpretation.)

```{r s3 cap ratings dataset}
d3_cap_ratings <- d3 %>%
  mutate(target_year = target_num/12) %>%
  mutate(factor = recode_factor(domain,
                                "BOD" = "Bodily sensations",
                                "NEG" = "Negative affect",
                                "POS" = "Social connection",
                                "COG" = "Cognition and control")) %>%
  select(-domain)
```

```{r s3 cap ratings summary stats}
d3_cap_ratings_sum <- d3_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup() %>%
  arrange(target_year, factor)

d3_cap_ratings_sum %>%
  mutate_if(is.numeric, funs(round_x(., 2)))
```

```{r s3 gam dataset}
d3_gam <- d3_cap_ratings %>%
  # implement correction for beta distribution
  mutate(response = case_when(response/100 < 0.005 ~ 0.005,
                              response/100 > 0.995 ~ 0.995,
                              TRUE ~ response/100))

contrasts(d3_gam$factor) <- cbind(bod_cog = c(1, 0, 0, 0),
                                  neg_cog = c(0, 1, 0, 0),
                                  soc_cog = c(0, 0, 1, 0))
```

```{r s3 gam}
if (!file.exists("../outputs/s3_gam.RDS")) {
  # takes ~5 minutes to run
  s3_gam <- gamm(response ~ s(target_year, by = factor) + factor, 
                 random = list(ResponseId = ~1,
                               capacity = ~1),
                 data = d3_gam,
                 family = betar,
                 method = "REML",
                 niterPQL = 40)

  saveRDS(s3_gam, "../outputs/s3_gam.RDS")

} else {
  s3_gam <- readRDS("../outputs/s3_gam.RDS")
}

summary(s3_gam$gam)
```

```{r tabs3gam}
tabs3gam <- model_parameters(s3_gam$gam)
```

```{r tabs3gamfixed, include = T}
tabs3gam %>% 
  filter(Component == "conditional") %>%
  select(-c(df, df_error, Component)) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Cognition and control (intercept)",
    grepl("bod", Parameter) ~ "Bodily sensations vs. Cognition and control",
    grepl("neg", Parameter) ~ "Negative affect vs. Cognition and control",
    grepl("soc", Parameter) ~ "Social connection vs. Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "Study 3, GAM results: Parametric coefficients (fixed effects).",
        digits = 2, align = c("lrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r tabs3gamsmooth, include = T}
tabs3gam %>% 
  filter(Component == "smooth_terms") %>%
  select(-c(Coefficient, SE, starts_with("CI"), df_error, Component)) %>%
  mutate(Parameter = case_when(
    grepl("Bod", Parameter) ~ "Target year: Bodily sensations",
    grepl("Neg", Parameter) ~ "Target year: Negative affect",
    grepl("Soc", Parameter) ~ "Target year: Social connection",
    grepl("Cog", Parameter) ~ "Target year: Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  rename(`Estimated df` = df) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "Study 3, GAM results: Approximate significance of smooth terms.",
        digits = 2, align = c("lrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r s3 gam predictions}
d3_gam_r1_pred <- d3_gam %>%
  mutate(pred = predict.gam(s3_gam$gam, type = "response", 
                            newdata = d3_gam)) %>%
  group_by(ResponseId, factor, target_year) %>%
  summarise(score = mean(response, na.rm = T),
            pred_score = mean(score, na.rm = T)) %>%
  ungroup()
```

```{r fig02b, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 3). Scores were calculated by averaging a participant’s responses to the two capacity items for each domain of mental life. Lighter lines represent individual participants’ scores, black points correspond to mean scores across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive model (beta regression).", fig.width = 4, fig.asp = 0.5}
d3_gam_r1_rsq <- round_x(summary(s3_gam$gam)$r.sq, 2)

fig02b <- d3_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            # color = "#160072",
            alpha = 0.2, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(data = . %>%
                    group_by(factor, target_year) %>%
                    multi_boot_standard(col = "score") %>%
                    ungroup,
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black",
                  # color = "#160072",
                  fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response\n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d3_gam_r1_rsq)))) +
  guides(color = guide_none())

# # save plot
# if (resave_plots == T | !file.exists("../outputs/fig02b.pdf")) {
#   ggsave("../outputs/fig02b.svg")
#   ggsave("../outputs/fig02b.pdf")
# }
```

```{r fig02, fig.cap="Perceived developmental trajectories for four domains of mental life (Studies 2-3). Lighter lines represent individual participants’ responses, black points correspond to mean responses across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive models (beta regressions). In Study 2 (Panel A), participants assessed 5 capacities within each domain, and assessed all capacities for a given target age before moving on to the next target age. In Study 3 (Panel B), participants assessed 2 capacities within each domain, and assessed a single capacity for all target ages before moving on to the next capacity.", fig.width = 4, fig.asp = 1.2}
plot_grid(fig02a, fig02b,
          ncol = 1, labels = "AUTO")

# save plot
if (resave_plots == T | !file.exists("../outputs/fig02.pdf")) {
  ggsave("../outputs/fig02.svg")
  ggsave("../outputs/fig02.pdf")
}
```

### Developmental mechanisms: Ratings

As a first step in assessing the lay theories underlying these perceptions of the development of human mental life, we asked participants to assess the importance of a variety of possible developmental mechanisms for each of the capacities included in this study. We assessed a wide range of potential developmental mechanisms for children’s capacities. We applied a variety of dimensionality reduction approaches in order to sort these developmental mechanisms into meaningful categories, and then modeled participants' ratings of the importance of these mechanism as a function of domain (_bodily sensations_, _negative affect_, _social connection_, and _cognition and control_).

```{r s3 mech ratings datasets a}
d3_mech_ratings_wide <- d3_dev_factor_rating %>%
  select(-domain) %>%
  spread(dev_factor, response) %>%
  unite(RepsonseId_capacity, c(ResponseId, capacity)) %>%
  column_to_rownames("RepsonseId_capacity")
```

#### Dimensionality reduction

Here we present two dimensionality reduction approaches: EFA via parallel analysis (analogous to the EFAs of mental capacity attributions from Studies 1-2) and hierarchical clustering.

##### EFA (via parallel analysis)

```{r s3 parallel analysis}
set.seed(94305) # this does not appear to work well -- may need to try a few times
s3_parallel <- fa.parallel(d3_mech_ratings_wide, plot = F)
```

```{r s3 efa parallel analysis}
s3_efa_par <- fa(d3_mech_ratings_wide, 
                 nfactors = s3_parallel$nfact, 
                 rotate = "oblimin")
# s3_efa_par %>% fa.sort()
```

Parallel analysis suggested retaining `r s3_parallel$nfact` factors, which we labeled _extrinsic (passive)_, _extrinsic (active)_, _intrinsic (after birth)_, and _intrinsic (before birth)_. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs3efaload); the amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs3efavacc).

```{r tabs3efaload, include = T, warning = F, message = F}
tabs3efaload <- model_parameters(s3_efa_par, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate_at(vars(starts_with("MR")),
            funs("bold" = ifelse(. >= 0.6, T, F))) %>%
  rename(`Extrinsic (passive)` = MR1,
         `Extrinsic (active)` = MR3,
         `Intrinsic (after birth)` = MR4,
         `Intrinsic (before birth)` = MR2)

tabs3efaload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 3, factor solution suggested by parallel analysis: Factor loadings, complexity, and uniqueness.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1.8cm", bold = tabs3efaload[,s3_parallel$nfact+4]) %>%
  column_spec(3, width = "1.8cm", bold = tabs3efaload[,s3_parallel$nfact+5]) %>%
  column_spec(4, width = "1.8cm", bold = tabs3efaload[,s3_parallel$nfact+6]) %>%
  column_spec(5, width = "1.8cm", bold = tabs3efaload[,s3_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs3efavacc, include = T, warning = F, message = F}
s3_efa_par$Vaccounted %>% 
  data.frame() %>% 
  rename(`Extrinsic (passive)` = MR1,
         `Extrinsic (active)` = MR3,
         `Intrinsic (after birth)` = MR4,
         `Intrinsic (before birth)` = MR2) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 3, factor solution suggested by parallel analysis: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r figS06 code, fig.width = 4.5, fig.asp = 0.5}
heatmap_fun(s3_efa_par, 
            factor_names = c("Extrinsic (passive)",
                             "Intrinsic (before birth)",
                             "Extrinsic (active)",
                             "Intrinsic (after birth)"))

# save plot
if (resave_plots == T | !file.exists("../outputs/figS06.pdf")) {
  ggsave("../outputs/figS06.svg")
  ggsave("../outputs/figS06.pdf")
}
```

```{r figS06, fig.cap="Study 3, factor solution suggested by parallel analysis."}
# knitr::include_graphics("../outputs/figS06.pdf")
```

##### Hierarchical clustering

In addition to the EFA just described, we also conducted a hierarchical cluster analysis, using Euclidean distance and Ward's agglomeration method. The resulting dendrogram is presented in Supplementary Figure \@ref(fig:figS07). We view this as providing further support for the idea that these mechanisms can be sorted into two high-level categories: "extrinsic" mechanisms (left) and "intrinsic" mechanisms (right).

```{r figS07 code, fig.width = 3, fig.asp = 0.6}
d3_mech_ratings_wide %>% 
  rename_with(~recode(., 
                      "preprogrammed" = "biologically ‘preprogrammed’",
                      "womb_experiences" = "experiences in the womb",
                      "body_grows" = "body grows and matures",
                      "brain_changes" = "brain changes",
                      "senses_improve" = "senses improve",
                      "observes_objects" = "observes objects and physical world",
                      "experiments" = "actively experiments",
                      "observes_people" = "observes people",
                      "interacts_people" = "interacts with people",
                      "people_teach" = "people explicitly teach")) %>%
  remove_missing() %>% 
  as.matrix() %>% 
  t() %>%
  dist(method = "euclidean") %>%
  hclust(method = "ward") %>%
  ggdendrogram(rotate = F) +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin = unit(c(5.5, 5.5, 5.5, 40), "points"))

# save plot
if (resave_plots == T | !file.exists("../outputs/figS07.pdf")) {
  ggsave("../outputs/figS07.svg")
  ggsave("../outputs/figS07.pdf")
}
```

```{r figS07, fig.cap="Hierarchical cluster analysis of importance ratings of developmental mechanisms (Study 3).", include = T}
knitr::include_graphics("../outputs/figS07.pdf")
```

#### Regression models

To assess which developmental mechanisms were considered to be the most or least important in the development of each of the individual mental capacities included in this study (*bodily sensations*, *negative affect*, *social connection*, and *cognition and control*), we conducted a multilevel linear regression regressing responses to this question onto type of developmental mechanism (extrinsic vs. intrinsic), domain, and an interaction between them, including the maximal random effects structure that allowed the model to converge. Using the "lme4" package for R [@bates.etal2015], we specified this model as follows: lmer(response ~ mechanism type * domain + (1 + mechanism type + domain | participant) + (1 | developmental mechanism) + (1 | capacity). We also conducted separate, analogous regression analyses for each domain separately. In these analyses, mechanism type and domain were effect-coded for comparisons to the grand mean collapsing across mechanism types and domains. (Note that these analyses omitted the mechanism "the child’s brain changes [...]", because it was the only mechanism that was difficult to classify as either intrinsic or extrinsic via EFA and clustering analyses; see previous section.)

The full results of these models are presented in Supplementary Tables \@ref(tab:tabs3lmerall) (all domains), \@ref(tab:tabs3lmerbod) (*bodily sensations*), \@ref(tab:tabs3lmerneg) (*negative affect*), \@ref(tab:tabs3lmersoc) (*social connection*), and \@ref(tab:tabs3lmercog) (*cognition and control*); see also Figure 3 (main text).

```{r s3 mech ratings datasets b}
d3_mech_ratings <- d3_dev_factor_rating %>%
  mutate(factor = recode_factor(domain,
                                "BOD" = "Bodily sensations",
                                "NEG" = "Negative affect",
                                "POS" = "Social connection",
                                "COG" = "Cognition and control")) %>%
  select(-domain) %>%
  mutate(dev_factor_cluster = case_when(
    dev_factor %in% c("senses_improve", "body_grows", 
                      "womb_experiences", "preprogrammed") ~ "intrinsic",
    dev_factor %in% c("interacts_people", "observes_people", "observes_objects", 
                      "people_teach", "experiments") ~ "extrinsic",
    dev_factor == "brain_changes" ~ NA_character_,
    TRUE ~ NA_character_
  )) %>%
  mutate(dev_factor_cluster = factor(dev_factor_cluster,
                                     levels = c("intrinsic", "extrinsic")))

contrasts(d3_mech_ratings$factor) <- cbind(bod_gm = c(1, -1, 0, 0),
                                           soc_gm = c(0, -1, 1, 0),
                                           cog_gm = c(0, -1, 0, 1))

# contrasts(d3_mech_ratings$factor) <- cbind(bod_gm = c(1, 0, -1, 0),
#                                            neg_gm = c(0, 1, -1, 0),
#                                            cog_gm = c(0, 0, -1, 1))

contrasts(d3_mech_ratings$dev_factor_cluster) <- cbind(ext_gm = c(-1, 1))
```

```{r s3 mech ratings summary stats}
# overall
d3_mech_ratings_sum_a <- d3_mech_ratings %>%
  filter(!is.na(dev_factor_cluster)) %>%
  # participant-level mean of means for each cluster
  group_by(ResponseId, dev_factor_cluster) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means for each cluster 
  group_by(dev_factor_cluster) %>%
  summarise(median = median(score_a, na.rm = T),
            mean = mean(score_a, na.rm = T),
            sd = sd(score_a, na.rm = T)) %>%
  ungroup() %>%
  arrange(dev_factor_cluster)

d3_mech_ratings_sum_a %>%
  mutate_if(is.numeric, funs(round_x(., 2)))

# by domain
d3_mech_ratings_sum_b <- d3_mech_ratings %>%
  filter(!is.na(dev_factor_cluster)) %>%
  # participant-level mean for each mechanism, for each domain
  group_by(ResponseId, dev_factor_cluster, dev_factor, factor) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # participant-level mean of means for each cluster, for each domain
  group_by(ResponseId, dev_factor_cluster, factor) %>%
  summarise(score_b = mean(score_a, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means of means for each domain 
  group_by(dev_factor_cluster, factor) %>%
  summarise(median = median(score_b, na.rm = T),
            mean = mean(score_b, na.rm = T),
            sd = sd(score_b, na.rm = T)) %>%
  ungroup() %>%
  arrange(factor, dev_factor_cluster)

d3_mech_ratings_sum_b %>%
  mutate_if(is.numeric, funs(round_x(., 2)))

# mean DIFFERENCES by domain
d3_mech_ratings_sum_c <- d3_mech_ratings %>%
  filter(!is.na(dev_factor_cluster)) %>%
  # participant-level mean for each cluster, for each capacity
  group_by(ResponseId, dev_factor_cluster, factor, capacity) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # participant-level cluster differences for each capacity
  spread(dev_factor_cluster, score_a) %>%
  mutate(diff_a = extrinsic - intrinsic) %>%
  select(-extrinsic, -intrinsic) %>%
  # participant-level mean differences, for each domain
  group_by(ResponseId, factor) %>%
  summarise(diff_b = mean(diff_a, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means of mean differences, for each domain 
  group_by(factor) %>%
  summarise(median = median(diff_b, na.rm = T),
            mean = mean(diff_b, na.rm = T),
            sd = sd(diff_b, na.rm = T)) %>%
  ungroup() %>%
  arrange(factor)

d3_mech_ratings_sum_c %>%
  mutate_if(is.numeric, funs(round_x(., 2)))
```

##### All domains

In brief, this analysis suggested that, collapsing across domains, intrinsic and extrinsic mechanisms were rated as equally important, but that the difference between intrinsic and extrinsic mechanisms varied across domains (see interaction terms).

```{r s3 mech ratings lmer}
s3_lmer <- lmer(response ~ dev_factor_cluster * factor + 
                  (1 + dev_factor_cluster + factor | ResponseId) + 
                  (1 | dev_factor) +
                  (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings)

summary(s3_lmer)
# model_parameters(s3_lmer)
```

```{r tabs3lmerall, include = T, message = F, warning = F}
model_parameters(s3_lmer, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factor_clusterext_gm", "Mechanism type", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (all domains): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Bodily sensations

For capacities in the domain of _bodily sensations_, this analysis suggested that participants rated intrinsic mechanisms as more important drivers of development than extrinsic mechanisms; see also Figure 3, panels A-B (main text).

```{r s3 mech ratings lmer bod}
s3_lmer_bod <- lmer(response ~ dev_factor_cluster + 
                      (1 + dev_factor_cluster | ResponseId) + 
                      (1 | dev_factor) + 
                      (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings %>% 
                  filter(factor == "Bodily sensations"))

summary(s3_lmer_bod)
# model_parameters(s3_lmer)
```

```{r tabs3lmerbod, include = T, message = F, warning = F}
model_parameters(s3_lmer_bod, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factor_clusterext_gm", "Mechanism type", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (bodily sensations): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Negative affect

For capacities in the domain of _negative affect_, this analysis suggested that participants rated intrinsic and extrinsic mechanisms as equally important drivers of development; see also Figure 3, panels A-B (main text).

```{r s3 mech ratings lmer neg}
s3_lmer_neg <- lmer(response ~ dev_factor_cluster + 
                      (1 + dev_factor_cluster | ResponseId) + 
                      (1 | dev_factor) + 
                      (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings %>% 
                  filter(factor == "Negative affect"))

summary(s3_lmer_neg)
# model_parameters(s3_lmer)
```

```{r tabs3lmerneg, include = T, message = F, warning = F}
model_parameters(s3_lmer_neg, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factor_clusterext_gm", "Mechanism type", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (negative affect): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Social connection

For capacities in the domain of _social connection_, this analysis suggested that participants rated extrinsic mechanisms as more important drivers of development than intrinsic mechanisms; see also Figure 3, panels A-B (main text).

```{r s3 mech ratings lmer soc}
s3_lmer_soc <- lmer(response ~ dev_factor_cluster + 
                      (1 + dev_factor_cluster | ResponseId) + 
                      (1 | dev_factor) + 
                      (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings %>% 
                  filter(factor == "Social connection"))

summary(s3_lmer_soc)
# model_parameters(s3_lmer)
```

```{r tabs3lmersoc, include = T, message = F, warning = F}
model_parameters(s3_lmer_soc, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factor_clusterext_gm", "Mechanism type", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (social connection): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Cognition and control

For capacities in the domain of _cognition and control_, this analysis suggested that participants rated extrinsic mechanisms as much more important drivers of development than intrinsic mechanisms; see also Figure 3, panels A-B (main text).

```{r s3 mech ratings lmer cog}
s3_lmer_cog <- lmer(response ~ dev_factor_cluster + 
                      (1 + dev_factor_cluster | ResponseId) + 
                      (1 | dev_factor) + 
                      (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings %>% 
                  filter(factor == "Cognition and control"))

summary(s3_lmer_cog)
# model_parameters(s3_lmer)
```

```{r tabs3lmercog, include = T, message = F, warning = F}
model_parameters(s3_lmer_cog, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factor_clusterext_gm", "Mechanism type", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (cognition and control): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r fig03a, fig.width = 4.5, fig.asp = 1.5}
fig03a <- d3_mech_ratings %>%
  mutate(dev_factor_cluster = case_when(
    is.na(dev_factor_cluster) ~ "Other",
    dev_factor_cluster == "intrinsic" ~ "INTRINSIC mechanisms",
    dev_factor_cluster == "extrinsic" ~ "EXTRINSIC mechanisms"),
    dev_factor_cluster = factor(dev_factor_cluster,
                                levels = c("INTRINSIC mechanisms", 
                                           "EXTRINSIC mechanisms",
                                           "Other")),
    dev_factor_text = recode_factor(
      dev_factor,
      "preprogrammed" = "biologically ‘preprogrammed’",
      "womb_experiences" = "experiences in the womb",
      "body_grows" = "body grows and matures",
      "brain_changes" = "brain changes",
      "senses_improve" = "senses improve",
      "observes_objects" = "observes objects and physical world",
      "experiments" = "actively experiments",
      "observes_people" = "observes people",
      "interacts_people" = "interacts with people",
      "people_teach" = "people explicitly teach"),
    capacity = str_wrap(gsub("_", " ", capacity), 14)) %>%
  ggplot(aes(x = reorder(str_wrap(dev_factor_text, 20), 
                         as.numeric(dev_factor)), 
             y = response, color = factor)) +
  facet_grid(rows = vars(reorder(capacity, as.numeric(factor))),
             cols = vars(dev_factor_cluster),
             scales = "free", space = "free") +
  geom_hline(yintercept = 3, lty = 2, color = "darkred") +
  geom_jitter(alpha = 0.6, size = 0.02) +
  geom_pointrange(data = . %>%
                    group_by(factor, capacity, 
                             dev_factor_cluster, 
                             dev_factor, dev_factor_text) %>%
                    multi_boot_standard(col = "response") %>%
                    ungroup(),
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black", fatten = 2) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 6, 3), minor_breaks = seq(0, 6, 1)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Developmental mechanism",
       y = "Response (0 = not at all important, 6 = extremely important)",
       color = "Domain") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))
```

```{r fig03b, fig.width = 1.5, fig.asp = 6.75/1.5}
fig03b <- d3_mech_ratings %>%
  # filter(!is.na(dev_factor_cluster)) %>%
  filter(!is.na(dev_factor)) %>%
  mutate(dev_factor_cluster = case_when(
    dev_factor == "brain_changes" ~ "brain changes",
    dev_factor_cluster == "intrinsic" ~ "INTRINSIC",
    dev_factor_cluster == "extrinsic" ~ "EXTRINSIC"),
    dev_factor_cluster = factor(dev_factor_cluster,
                                levels = c("INTRINSIC", 
                                           "EXTRINSIC",
                                           "brain changes"))) %>%  
  # participant-level means for each cluster, by capacity
  group_by(ResponseId, dev_factor_cluster, factor, capacity) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # participant-level mean of means for each cluster, by domain
  group_by(ResponseId, dev_factor_cluster, factor) %>%
  summarise(score_b = mean(score_a, na.rm = T)) %>%
  ungroup() %>%
  mutate(blank_var = "  ") %>%
  ggplot(aes(x = reorder(str_pad(dev_factor_cluster, 
                                 width = 28, side = "left"),
                         as.numeric(dev_factor_cluster)), 
             y = score_b, color = factor)) +
  facet_grid(cols = vars(blank_var),
             rows = vars(reorder(toupper(factor), as.numeric(factor))),
             scales = "free", space = "free") +
  geom_hline(yintercept = 3, lty = 2, color = "darkred") +
  geom_path(data = . %>%
              filter(dev_factor_cluster != "brain changes"),
            aes(group = ResponseId), 
            alpha = 0.1, size = 0.02,
            position = position_jitter(width = 0.2, 
                                       height = 0, 
                                       seed = 94305)) +
  geom_point(alpha = 0.6, size = 0.02,
             position = position_jitter(width = 0.2, 
                                        height = 0, 
                                        seed = 94305)) +
  # group-level mean of means of means for each cluster, by domain
  geom_pointrange(data = . %>%
                    group_by(factor, dev_factor_cluster) %>%
                    multi_boot_standard(col = "score_b") %>%
                    ungroup(),
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black", fatten = 2) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 6, 3), minor_breaks = seq(0, 6, 1)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Developmental mechanism type",
       y = "Mean response (0 = not at all important, 6 = extremely important)",
       color = "Domain") +
  guides(color = guide_none())
  # guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))
# fig03b
```

### Developmental mechanisms: Selection of "most important"

As described in the main text, we also assessed participants' responses to forced-choice questions about the "most important" factor driving the development of each of these capacities. In the main text, we presented a visualization of these choices (Figure 3, Panel C). Here, we present multilevel generalized linear model (logistic regressions) to predict the likelihood of selecting extrinsic vs. intrinsic mechanisms by domain. Using the "lme4" package for R [@bates.etal2015], we specified this model as follows: glmer(choice of extrinsic mechanism ~ domain + (1 | participant) + (1 | capacity). We also conducted separate, analogous regression analyses for each domain separately. In these analyses, domain was effect-coded for comparisons to the grand mean collapsing across domains. (As in the previous section, these analyses omitted the mechanism "the child’s brain changes [...]", because it was the only mechanism that was difficult to classify as either intrinsic or extrinsic via EFA and clustering analyses.)

The full results of these models are presented in Supplementary Tables \@ref(tab:tabs3glmerall) (all domains), \@ref(tab:tabs3glmerbod) (*bodily sensations*), \@ref(tab:tabs3glmerneg) (*negative affect*), \@ref(tab:tabs3glmersoc) (*social connection*), and \@ref(tab:tabs3glmercog) (*cognition and control*); see also Figure 3 (main text).

```{r s3 mech choice datasets}
d3_mech_choice <- d3_dev_factor_most_important_choice %>%
  mutate(factor = recode_factor(domain,
                                "BOD" = "Bodily sensations",
                                "NEG" = "Negative affect",
                                "POS" = "Social connection",
                                "COG" = "Cognition and control"),
         dev_factor = recode_factor(
           response,
           "The child is biologically \"preprogrammed\" to have this ability" = "preprogrammed",
           "The child has experiences in the womb that give them this ability" = "womb_experiences",
           "The child's body grows and matures (for example, muscles get stronger, child gets taller)" = "body_grows",
           "The child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" = "brain_changes",
           "The child's senses improve (for example, vision gets sharper, hearing improves)" = "senses_improve",
           "The child observes the objects and the physical world around him or her" = "observes_objects",
           "The child actively experiments with how to do this" = "experiments",
           "The child observes the people around him or her" = "observes_people",
           "The child interacts with the people around him or her" = "interacts_people",
           "People explicitly teach the child how to do this" = "people_teach",
           .default = "other")) %>%
  select(-domain, -response) %>%
  mutate(dev_factor_cluster = case_when(
    dev_factor %in% c("senses_improve", "body_grows", 
                      "womb_experiences", "preprogrammed") ~ "intrinsic",
    dev_factor %in% c("interacts_people", "observes_people", "observes_objects", 
                      "people_teach", "experiments") ~ "extrinsic",
    dev_factor == "brain_changes" ~ NA_character_,
    TRUE ~ NA_character_
  )) %>%
  mutate(dev_factor_cluster = factor(dev_factor_cluster,
                                     levels = c("intrinsic", "extrinsic")))

contrasts(d3_mech_choice$factor) <- cbind(bod_gm = c(1, -1, 0, 0),
                                           soc_gm = c(0, -1, 1, 0),
                                           cog_gm = c(0, -1, 0, 1))

# contrasts(d3_mech_choice$factor) <- cbind(bod_gm = c(1, 0, -1, 0),
#                                            neg_gm = c(0, 1, -1, 0),
#                                            cog_gm = c(0, 0, -1, 1))

contrasts(d3_mech_choice$dev_factor_cluster) <- cbind(ext_gm = c(-1, 1))
```

```{r s3 mech choice summary stats}
# overall
d3_mech_choice_sum_a <- d3_mech_choice %>%
  count(dev_factor_cluster) %>%
  mutate(prop = n/sum(n))

d3_mech_choice_sum_a %>%
  mutate_at(vars(prop), funs(round_x(., 2)))

# by domain
d3_mech_choice_sum_b <- d3_mech_choice %>%
  count(factor, dev_factor_cluster) %>%
  group_by(factor) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()

d3_mech_choice_sum_b %>%
  mutate_at(vars(prop), funs(round_x(., 2)))
```

##### All domains

This analysis suggested that participants' choice of extrinsic vs. intrinsic mechanisms as the "most important" driver of development varied by domain (see interaction terms).

```{r s3 mech choice glmer}
s3_glmer <- glmer(extrinsic ~ factor + 
                    (1 | ResponseId) + 
                    (1 | capacity),
                  control = glmerControl(optimizer = "bobyqa"),
                  data = d3_mech_choice %>%
                    mutate(extrinsic = case_when(dev_factor_cluster == "extrinsic" ~ 1,
                                                 dev_factor_cluster == "intrinsic" ~ 0,
                                                 is.na(dev_factor_cluster) ~ NA_real_,
                                                 TRUE ~ NA_real_)),
                  family = "binomial")

summary(s3_glmer)
# model_parameters(s3_glmer)
```

```{r tabs3glmerall, include = T, message = F, warning = F}
model_parameters(s3_glmer, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel generalized linear model results (all domains): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Bodily sensastions

For capacities in the domain of _bodily sensations_, this analysis suggested that participants were much more likely to select intrinsic mechanisms than extrinsic mechanisms as the most important drivers of development; see also Figure 3, panel C (main text).

```{r s3 mech choice glmer bodily sensations}
s3_glmer_bod <- glmer(extrinsic ~ 1 + 
                        (1 | ResponseId) + 
                        (1 | capacity),
                      control = glmerControl(optimizer = "bobyqa"),
                      data = d3_mech_choice %>%
                        filter(factor == "Bodily sensations") %>%
                        mutate(extrinsic = case_when(dev_factor_cluster == "extrinsic" ~ 1,
                                                     dev_factor_cluster == "intrinsic" ~ 0,
                                                     is.na(dev_factor_cluster) ~ NA_real_,
                                                     TRUE ~ NA_real_)),
                      family = "binomial")

summary(s3_glmer_bod)
# model_parameters(s3_glmer)
```

```{r tabs3glmerbod, include = T, message = F, warning = F}
model_parameters(s3_glmer_bod, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel generalized linear model results (bodily sensations): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Negative affect

For capacities in the domain of _negative affect_, this analysis suggested that participants were equally likely to select intrinsic and extrinsic mechanisms as the most important drivers of development; see also Figure 3, panel C (main text).

```{r s3 mech choice glmer negative affect}
s3_glmer_neg <- glmer(extrinsic ~ 1 + 
                        (1 | ResponseId) + 
                        (1 | capacity),
                      control = glmerControl(optimizer = "bobyqa"),
                      data = d3_mech_choice %>%
                        filter(factor == "Negative affect") %>%
                        mutate(extrinsic = case_when(dev_factor_cluster == "extrinsic" ~ 1,
                                                     dev_factor_cluster == "intrinsic" ~ 0,
                                                     is.na(dev_factor_cluster) ~ NA_real_,
                                                     TRUE ~ NA_real_)),
                      family = "binomial")

summary(s3_glmer_neg)
# model_parameters(s3_glmer)
```

```{r tabs3glmerneg, include = T, message = F, warning = F}
model_parameters(s3_glmer_neg, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel generalized linear model results (negative affect): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Social connection

For capacities in the domain of _social connection_, this analysis suggested that participants were equally likely to select intrinsic and extrinsic mechanisms as the most important drivers of development; see also Figure 3, panel C (main text).

```{r s3 mech choice glmer social connection}
s3_glmer_soc <- glmer(extrinsic ~ 1 + 
                        (1 | ResponseId) + 
                        (1 | capacity),
                      control = glmerControl(optimizer = "bobyqa"),
                      data = d3_mech_choice %>%
                        filter(factor == "Social connection") %>%
                        mutate(extrinsic = case_when(dev_factor_cluster == "extrinsic" ~ 1,
                                                     dev_factor_cluster == "intrinsic" ~ 0,
                                                     is.na(dev_factor_cluster) ~ NA_real_,
                                                     TRUE ~ NA_real_)),
                      family = "binomial")

summary(s3_glmer_soc)
# model_parameters(s3_glmer)
```

```{r tabs3glmersoc, include = T, message = F, warning = F}
model_parameters(s3_glmer_soc, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel generalized linear model results (social connection): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Cognition and control

For capacities in the domain of _cognition and control_, this analysis suggested that participants were much more likely to select extrinsic mechanisms than intrinsic mechanisms as the most important drivers of development; see also Figure 3, panel C (main text).

```{r s3 mech choice glmer cognition and control}
s3_glmer_cog <- glmer(extrinsic ~ 1 + 
                        (1 | ResponseId) + 
                        (1 | capacity),
                      control = glmerControl(optimizer = "bobyqa"),
                      data = d3_mech_choice %>%
                        filter(factor == "Cognition and control") %>%
                        mutate(extrinsic = case_when(dev_factor_cluster == "extrinsic" ~ 1,
                                                     dev_factor_cluster == "intrinsic" ~ 0,
                                                     is.na(dev_factor_cluster) ~ NA_real_,
                                                     TRUE ~ NA_real_)),
                      family = "binomial")

summary(s3_glmer_cog)
# model_parameters(s3_glmer)
```

```{r tabs3glmercog, include = T, message = F, warning = F}
model_parameters(s3_glmer_cog, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensations (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel generalized linear model results (cognition and control): Fixed effects.",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r figS08a, fig.width = 4.5, fig.asp = 1.5}
figS08a <- d3_mech_choice %>%
  mutate(dev_factor_cluster = case_when(
    dev_factor_cluster == "intrinsic" ~ "INTRINSIC mechanisms",
    is.na(dev_factor_cluster) ~ "Other",
    dev_factor_cluster == "extrinsic" ~ "EXTRINSIC mechanisms"),
    dev_factor_cluster = factor(dev_factor_cluster,
                                levels = c("INTRINSIC mechanisms", 
                                           # "Other",
                                           "EXTRINSIC mechanisms")),
    dev_factor_text = recode_factor(
      dev_factor,
      "preprogrammed" = "biologically ‘preprogrammed’",
      "womb_experiences" = "experiences in the womb",
      "body_grows" = "body grows and matures",
      "brain_changes" = "brain changes",
      "senses_improve" = "senses improve",
      "observes_objects" = "observes objects and physical world",
      "experiments" = "actively experiments",
      "observes_people" = "observes people",
      "interacts_people" = "interacts with people",
      "people_teach" = "people explicitly teach"),
    capacity = str_wrap(gsub("_", " ", capacity), 14)) %>%
  ggplot(aes(x = reorder(str_wrap(dev_factor_text, 20), 
                         as.numeric(dev_factor)), 
             fill = factor, color = factor)) +
  facet_grid(rows = vars(reorder(capacity, as.numeric(factor))),
             cols = vars(dev_factor_cluster),
             scales = "free_x", space = "free_x") +
  geom_bar(stat = "count", alpha = 0.6) +
  scale_fill_viridis_d(end = 0.9, option = "viridis", 
                       aesthetics = c("color", "fill")) +
  scale_y_continuous(breaks = seq(0, 
                                  d3_mech_choice %>% 
                                    distinct(ResponseId) %>% 
                                    count() %>% 
                                    pull(n), 
                                  100),
                     limits = c(0, 
                                d3_mech_choice %>% 
                                  distinct(ResponseId) %>% 
                                  count() %>% 
                                  pull(n))) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Developmental mechanism",
       y = "Count of responses",
       color = "Domain", fill = "Domain") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)),
         fill = guide_legend(override.aes = list(alpha = 1, size = 2)))
```

```{r figS08b, fig.width = 1.5, fig.asp = 6.75/1.5}
figS08b <- d3_mech_choice %>%
  filter(!is.na(dev_factor_cluster)) %>%
  count(dev_factor_cluster, factor) %>%
  mutate(dev_factor_cluster = case_when(
    # is.na(dev_factor_cluster) ~ "Other",
    dev_factor_cluster == "intrinsic" ~ "INTRINSIC",
    dev_factor_cluster == "extrinsic" ~ "EXTRINSIC"),
    dev_factor_cluster = factor(dev_factor_cluster,
                                levels = c("INTRINSIC", 
                                           "EXTRINSIC")),
                                           # "Other")),
    blank_var = "  ") %>%
  ggplot(aes(x = reorder(str_pad(dev_factor_cluster, 
                                 width = 28, side = "left"),
                         as.numeric(dev_factor_cluster)),
             y = n,
             fill = factor, color = factor)) +
  facet_grid(cols = vars(blank_var),
             rows = vars(reorder(toupper(factor), as.numeric(factor))),
             scales = "free_x", space = "free_x") +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_viridis_d(end = 0.9, option = "viridis", aesthetics = c("color", "fill")) +
  scale_y_continuous(breaks = seq(0, 
                                  2 * d3_mech_choice %>% 
                                    distinct(ResponseId) %>% 
                                    count() %>% 
                                    pull(n), 
                                  2 * 100),
                     limits = c(0, 
                                2 * d3_mech_choice %>% 
                                  distinct(ResponseId) %>% 
                                  count() %>% 
                                  pull(n))) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Developmental mechanism type",
       y = "Count of responses",
       color = "Domain", fill = "Domain") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)),
         fill = guide_legend(override.aes = list(alpha = 1, size = 2)))
# figS08b
```

```{r figS08, fig.cap="XX", fig.width = 4.5, fig.asp = 6.75/6}
plot_grid(figS08a + guides(fill = guide_none(), color = guide_none()),
          figS08b + guides(fill = guide_none(), color = guide_none()) + labs(y = NULL), 
          nrow = 1, rel_widths = c(4.5, 1.5), labels = "AUTO")

# save plot
if (resave_plots == T | !file.exists("../outputs/figS08.pdf")) {
  ggsave("../outputs/figS08.svg")
  ggsave("../outputs/figS08.pdf")
}
```

```{r fig03c, fig.width = 1.5, fig.asp = 6.75/1.5}
fig03c <- d3_mech_choice %>%
  # filter(!is.na(dev_factor_cluster)) %>%
  filter(dev_factor != "other") %>%
  mutate(dev_factor_cluster = case_when(
    dev_factor == "brain_changes" ~ "brain changes",
    dev_factor_cluster == "intrinsic" ~ "INTRINSIC",
    dev_factor_cluster == "extrinsic" ~ "EXTRINSIC"),
    dev_factor_cluster = factor(dev_factor_cluster,
                                levels = c("INTRINSIC", 
                                           "EXTRINSIC",
                                           "brain changes"))) %>%
  count(dev_factor_cluster, factor) %>%
  mutate(blank_var = "  ") %>%
  ggplot(aes(x = reorder(str_pad(dev_factor_cluster, 
                                 width = 28, side = "left"),
                         as.numeric(dev_factor_cluster)),
             y = n,
             fill = factor, color = factor)) +
  facet_grid(cols = vars(blank_var),
             rows = vars(reorder(toupper(factor), as.numeric(factor))),
             scales = "free_x", space = "free_x") +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_viridis_d(end = 0.9, option = "viridis", aesthetics = c("color", "fill")) +
  scale_y_continuous(breaks = seq(0, 
                                  2 * d3_mech_choice %>% 
                                    distinct(ResponseId) %>% 
                                    count() %>% 
                                    pull(n), 
                                  2 * 100),
                     limits = c(0, 
                                2 * d3_mech_choice %>% 
                                  distinct(ResponseId) %>% 
                                  count() %>% 
                                  pull(n))) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Developmental mechanism type",
       y = "Count of responses",
       color = "Domain", fill = "Domain") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)),
         fill = guide_legend(override.aes = list(alpha = 1, size = 2)))
# fig03c
```

```{r fig03, fig.cap="Perceived importance of various mechanisms in the development of four domains of mental life (Study 3); see main text for the full text of each mechanism. Panel A shows ratings for each developmental mechanism and both of the capacities within each domain; Panel B shows mean ratings for extrinsic vs. intrinsic mechanisms for each domain of capacities; and Panel C shows the percentage of trials on which participants selected extrinsic vs. intrinsic mechanisms as the 'most important' driver of development. Lighter points and lines represent individual participants’ responses, black points correspond to mean scores across the sample, and error bars are bootstrapped 95% confidence intervals. The dotted red line at the midpoint of the response scale in Panels A and B is intended to aid visual comparison across domains.", fig.width = 5, fig.asp = 6.75/(4.5 + 1.5 + 1.5)}
plot_grid(fig03a + guides(color = guide_none()),
          NULL, # to adjust spacing between panels
          fig03b, # + labs(x = " "),
          NULL,
          fig03c + 
            labs(x = " ",
                 y = "Selection of 'most important' mechanism") +
            scale_y_continuous(limits = c(0, 2*nrow(d3_demo)),
                               breaks = c(0, nrow(d3_demo), 2*nrow(d3_demo)),
                               labels = c("0%", "50%", "100%")) +
            guides(fill = guide_none(),
                   color = guide_none()), 
          nrow = 1, rel_widths = c(4.5, 0.1, 1.5, 0.1, 1.5), 
          labels = c("A", "", "B", "", "C"))

# save plot
if (resave_plots == T | !file.exists("../outputs/fig03.pdf")) {
  ggsave("../outputs/fig03.svg")
  ggsave("../outputs/fig03.pdf")
}
```

```{r s3 export results}
# save numerical results
s3_results <- list(demo = d3_demo,
                   cap_rat_ss = d3_cap_ratings_sum,
                   mech_rat_ss_a = d3_mech_ratings_sum_a,
                   mech_rat_ss_b = d3_mech_ratings_sum_b,
                   mech_rat_ss_c = d3_mech_ratings_sum_c,
                   gam = s3_gam, 
                   lmer = s3_lmer, 
                   lmer_bod = s3_lmer_bod,
                   lmer_neg = s3_lmer_neg,
                   lmer_soc = s3_lmer_soc,
                   lmer_cog = s3_lmer_cog,
                   mech_choice_ss_a = d3_mech_choice_sum_a,
                   mech_choice_ss_b = d3_mech_choice_sum_b,
                   glmer = s3_glmer, 
                   glmer_bod = s3_glmer_bod,
                   glmer_neg = s3_glmer_neg,
                   glmer_soc = s3_glmer_soc,
                   glmer_cog = s3_glmer_cog)

saveRDS(s3_results, "../outputs/s3_results.rds")
```

```{r}
# ADDITIONAL EXPLORATORY ANALYSES  NOT DESCRIBED IN SUPPLEMENT
```

```{r figS09a, fig.width = 4.5, fig.asp = 0.4}
figS09a <- d2_gam %>% 
  group_by(factor, capacity, target_year) %>%
  summarise(var = var(response, na.rm = T)) %>%
  ungroup() %>%
  ggplot(aes(x = target_year, y = var, color = factor)) +
  facet_grid(cols = vars(factor)) +
  geom_path(aes(group = capacity), 
            alpha = 0.1, size = 0.5,
            position = position_jitter(width = 0.2, 
                                       height = 0, 
                                       seed = 94305)) +
  geom_point(alpha = 1, size = 0.5,
             position = position_jitter(width = 0.2, 
                                        height = 0, 
                                        seed = 94305)) +
  geom_text(data = . %>% 
              filter(target_year == 5) %>%
              group_by(factor) %>%
              filter(var == max(var) | var == min(var)) %>%
              ungroup(),
            aes(x = 5.2, label = str_wrap(gsub("_", " ", capacity), 18)),
            color = "black", hjust = 0, size = 2) +
  geom_pointrange(data = . %>%
                    group_by(factor) %>%
                    multi_boot_standard(col = "var") %>%
                    ungroup(),
                  aes(x = 2.5, y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black", shape = 15, size = 0.5) +
  geom_pointrange(data = . %>%
                    filter(target_year == 0) %>%
                    group_by(factor) %>%
                    multi_boot_standard(col = "var") %>%
                    ungroup(),
                  aes(x = -0.5, y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "#f94802", shape = 15, size = 0.3) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_x_continuous(limits = c(NA, 9), breaks = seq(0, 5, 1)) +
  scale_y_continuous(limits = c(0, 0.16)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Variance") +
  guides(color = guide_none())
```

```{r figS09b, fig.width = 4.5, fig.asp = 0.4}
figS09b <- d3_gam %>% 
  group_by(factor, capacity, target_year) %>%
  summarise(var = var(response, na.rm = T)) %>%
  ungroup() %>%
  ggplot(aes(x = target_year, y = var, color = factor)) +
  facet_grid(cols = vars(factor)) +
  geom_path(aes(group = capacity), 
            alpha = 0.1, size = 0.5,
            position = position_jitter(width = 0.2, 
                                       height = 0, 
                                       seed = 94305)) +
  geom_point(alpha = 1, size = 0.5,
             position = position_jitter(width = 0.2, 
                                        height = 0, 
                                        seed = 94305)) +
  geom_text(data = . %>% filter(target_year == 5),
            aes(x = 5.2, label = str_wrap(gsub("_", " ", capacity), 18)),
            color = "black", hjust = 0, size = 2) +
  geom_pointrange(data = . %>%
                    group_by(factor) %>%
                    multi_boot_standard(col = "var") %>%
                    ungroup(),
                  aes(x = 2.5, y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "black", shape = 15, size = 0.5) +
  geom_pointrange(data = . %>%
                    filter(target_year == 0) %>%
                    group_by(factor) %>%
                    multi_boot_standard(col = "var") %>%
                    ungroup(),
                  aes(x = -0.5, y = mean, ymin = ci_lower, ymax = ci_upper),
                  color = "#f94802", shape = 15, size = 0.3) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_x_continuous(limits = c(NA, 9), breaks = seq(0, 5, 1)) + 
  scale_y_continuous(limits = c(0, 0.16)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Variance") +
  guides(color = guide_none())
```

```{r figS09 code, fig.width = 4, fig.asp = 1.2}
plot_grid(figS09a, figS09b,
          ncol = 1, labels = "AUTO")

# save plot
if (resave_plots == T | !file.exists("../outputs/figS09.pdf")) {
  ggsave("../outputs/figS09.svg")
  ggsave("../outputs/figS09.pdf")
}
```

```{r figS09, fig.cap="Variance of capacity attributions in four domains of mental life (Studies 2-3). Lighter dots and lines represent variance across participnts for a particular capacity at a particular target age, red points correpond to mean variance across capacities at birth, black points correspond to mean variance across capacities and across all target ages, and error bars are bootstrapped 95% confidence intervals. In Study 2 (Panel A), participants assessed 5 capacities within each domain, and assessed all capacities for a given target age before moving on to the next target age. In Study 3 (Panel B), participants assessed 2 capacities within each domain, and assessed a single capacity for all target ages before moving on to the next capacity."}
knitr::include_graphics("../outputs/figS09.pdf")
```

# Selection of capacities: Comparison to previous studies

Supplementary Table \@ref(tab:tabs1efaparload) provides an item-by-item comparison of the capacities included in this study to capacities included in previous studies of mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

```{r capacities comparison df}
d_cap_comp <- read_csv("../data/capacities_comparison.csv") %>%
  mutate_if(is.character, funs(tolower(.))) %>%
  mutate_if(is.character, funs(gsub("recognizing others.", "recognizing others'", .))) %>%
  mutate_if(is.character, funs(gsub("self.control", "self-control", .))) %>%
  left_join(d1_scales %>%
              mutate(capacity = gsub("_", " ", capacity),
                     capacity = gsub("recognizing others ", "recognizing others' ", capacity),
                     capacity = gsub("self.control", "self-control", capacity)) %>%
              rename(`Study 1` = capacity)) %>%
  mutate_at(vars(`Weisman et al. (2017a)`, `Weisman et al. (2017b)`, 
                 `Gray et al. (2007)`, `Malle (2019, Study 1)`, `Malle (2019, Study 2)`),
            funs(case_when(is.na(`Study 1`) & grepl("similar to", .) ~ NA_character_, TRUE ~ .))) %>%
  arrange(order, `Weisman et al. (2017a)`, `Weisman et al. (2017b)`, 
          `Gray et al. (2007)`, `Malle (2019, Study 1)`, `Malle (2019, Study 2)`) %>%
  filter(!(is.na(`Study 1`) & 
             is.na(`Weisman et al. (2017a)`) &
             is.na(`Weisman et al. (2017b)`) &
             is.na(`Gray et al. (2007)`) &
             is.na(`Malle (2019, Study 1)`) & 
             is.na(`Malle (2019, Study 2)`))) %>%
  mutate_if(is.character, funs(replace_na(., "-"))) %>%
  mutate(factor = case_when(is.na(factor) ~ "Not included", TRUE ~ as.character(factor)),
         factor = factor(factor, 
                         levels = c("Bodily sensations", "Negative affect",
                                    "Social connection", "Cognition and control", 
                                    "Not included"))) %>%
  select(factor, everything(), -loading, -order) %>%
  rename(Domain = factor) %>%
  ungroup()
```

```{r tabcapcomp, include = T, warning = F, message = F}
d_cap_comp %>%
  select(-Domain) %>%
  mutate(`Study 1` = case_when(`Study 2` != "-" ~ paste0(`Study 1`, 
                                                         footnote_marker_symbol(1)),
                               TRUE ~ `Study 1`)) %>%
  mutate(`Study 1` = case_when(`Study 3` != "-" ~ paste0(`Study 1`, 
                                                         footnote_marker_symbol(2)),
                               TRUE ~ `Study 1`)) %>%
  select(-`Study 2`, -`Study 3`) %>%
  mutate_at(vars(`Gray et al. (2007)`),
            funs(gsub("\\..*", "", .))) %>%
  kable("latex", booktabs = T, escape = F, longtable = T) %>%
  kable_styling(latex_options = c("repeat_header", "striped"), font_size = 8) %>%
  add_header_above(c("Current studies" = 1,
                     "Previous studies" = 5)) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2:6, width = "2cm") %>%
  pack_rows(group_label = "Bodily sensations", start_row = 1, end_row = 11) %>%
  pack_rows("Negative affect", 12, 16) %>%
  pack_rows("Social connection", 17, 38) %>%
  pack_rows("Cognition and control", 39, 60) %>%
  pack_rows("Not included", 61, nrow(d_cap_comp)) %>%
  footnote(symbol = c("Also used in Study 2; ", "Also unsed in Study 3"),
           symbol_title = "Notes: ",
           footnote_as_chunk = T)
```

\newpage
# References