### 2024-12-30 paper/cover.tex
add cover pages and de-anonymize

\begin{titlepage}
    \begin{center}
        \vspace*{1in}
        {\Large Beliefs about the development of mental life} \\[2cm]

        Kara Weisman\\
        \textit{Department of Psychology, University of California, Riverside}\par
        \vspace{0.5cm}
        Lucy S. King\\
        \textit{Department of Psychology and Human Development, Vanderbilt University}\par
        \vspace{0.5cm}
        Kathryn Humphreys\\
        \textit{Department of Psychology and Human Development, Vanderbilt University}\par

    \end{center}
    \vfill
    \textbf{Author Note} \\
    \setlength{\parindent}{2em}
    \indent Kara Weisman: https://orcid.org/0000-0002-3706-7554. Lucy S. King: https://orcid.org/0000-0002-2552-0614. Kathryn Humphreys: https://orcid.org/0000-0002-5715-6597.

    \indent Study materials, preregistrations, data, analysis scripts, and links to the fully reproducible manuscript and supplemental materials are openly available via the project’s Open Science Framework page (https://osf.io/xrznd/; DOI: 10.17605/OSF.IO/XRZND). Funding was provided by the William R. and Sara Hart Kimball Stanford Graduate Fellowship (Weisman); the National Science Foundation (Weisman and King: Graduate Research Fellowships; Humphreys: CAREER 2042285), the National Institutes of Health (King: F32 HD105385), and the Jacobs Foundation (Humphreys: Early Career Research Fellowship 2017-1261-05).

    \indent Correspondence should be addressed to Kara Weisman at kgweisman@gmail.com.
\end{titlepage}