### 2025-01-06 paper/cover_anonymized.tex
make anonymized versions

\begin{titlepage}
    \begin{center}
        \vspace*{1in}
        {\Large Beliefs about the development of mental life} \\[2cm]

        [Author names omitted for anonymous review]

    \end{center}
    \vfill
    \textbf{Author Note} \\
    \setlength{\parindent}{2em}

    \indent Study materials, preregistrations, data, analysis scripts, and links to the fully reproducible manuscript and supplemental materials are openly available via the project’s Open Science Framework page (anonymized link for peer review: \url{https://osf.io/xrznd/?view_only=d9eb55e46028453cbd90534a003d3f52}). Funding sources and contact information omitted for anonymous review.

\end{titlepage}