### 2025-02-14 supplement/functional-PCA.Rmd

---
title: 'Baby mental life: fPCA'
author: "Kara Weisman"
date: "2025-02-03"
output: html_document
---

```{r setup, set.seed(94305), include = F}
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(include = F)
```

```{r load packages}
library(ggdendro)
library(dendextend)
library(psych)
library(lme4)
library(mgcv)
library(kableExtra)
library(langcog)
library(parameters)
library(cowplot)
library(fda)
library(tidyverse)
library(rsq)
library(ggnewscale)
```

```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", chdir = TRUE)))
```

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four core capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the intuitive theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

In this RMarkdown file, we apply a "functional PCA" approach to analyzing the data from Study 2.

```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
# suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", chdir = TRUE)))
# suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", chdir = TRUE)))
```

```{r}
d2_scales <- data.frame(
  capacity = colnames(d2_all)) %>%
  mutate(
    factor = case_when(
      capacity %in% c(
        "getting_hungry", "feeling_pain", "feeling_tired", "hearing_sounds",
        "feeling_physically_uncomfortable") ~ "Bodily sensation",
      capacity %in% c(
        "feeling_overwhelmed", "feeling_distressed", "feeling_frustrated",
        "feeling_helpless", "feeling_lonely") ~ "Negative affect",
      capacity %in% c(
        "feeling_happy", "finding_something_funny", "feeling_excited",
        "loving_somebody", "learning_from_other_people") ~ "Social connection",
      capacity %in% c(
        "having_self_control", "controlling_their_emotions", "telling_right_from_wrong",
        "planning", "reasoning_about_things") ~ "Cognition and control")) %>%
  mutate(factor = factor(
    factor, 
    levels = c(
      "Bodily sensation", "Negative affect", 
      "Social connection", "Cognition and control"))) %>%
  arrange(factor) %>%
  mutate(loading = NA,
         order = 1:nrow(.))
```

```{r}
d2_cap_ratings <- d2_all %>%
  rownames_to_column("ResponseId_target") %>%
  separate(ResponseId_target, c("ResponseId", "target")) %>%
  mutate(target_year = recode(
    target,
    target00mo = 0, # newborns
    target0Xmo = 4/365, # 4 days
    target01mo = 1/12, # 1 month
    target02mo = 2/12, 
    target04mo = 4/12, 
    target06mo = 6/12, 
    target09mo = 9/12,
    target12mo = 12/12,
    target18mo = 18/12,
    target24mo = 24/12,
    target36mo = 36/12,
    target48mo = 48/12,
    target60mo = 60/12)) %>%
  gather(capacity, response, -ResponseId, -starts_with("target")) %>%
  left_join(d2_scales %>% distinct(factor, capacity))
```

### Functional PCA

```{r functional_pca_study2_multivariate, echo=TRUE, message=FALSE, warning=FALSE}
# # Step 1: Reshape the data to wide format (each Capacity × Target Age as a column)
# d2_ratings_matrix <- d2_cap_ratings %>%
#   dplyr::select(-c(target, factor)) %>%
#   unite("capacity_age", capacity, target_year, sep = "_") %>%
#   pivot_wider(names_from = capacity_age, values_from = response) %>%
#   column_to_rownames("ResponseId") %>%
#   as.matrix()

# Step 1: Define the target ages based on 'target_year'
ages <- sort(unique(d2_cap_ratings$target_year))  # Numeric ages in years

# Step 2: Create a B-spline basis for smoothing
basis <- create.bspline.basis(rangeval = range(ages), nbasis = 4)  # Adjust nbasis if needed

# Step 3: Get the list of unique capacities
capacities <- unique(d2_cap_ratings$capacity)

# Step 4: Smooth the data for each capacity separately
smoothed_list <- lapply(capacities, function(cap) {
  # Subset data for the current capacity and reshape to wide format
  ratings_submatrix <- d2_cap_ratings %>%
    filter(capacity == cap) %>%
    dplyr::select(-factor, -target, -capacity) %>%
    pivot_wider(names_from = target_year, values_from = response) %>%
    column_to_rownames("ResponseId") %>%
    as.matrix()

  # Apply smoothing
  smooth.basis(argvals = ages, y = t(ratings_submatrix), fdParobj = basis)$fd
})

# # Step 5: Combine smoothed data for multivariate functional PCA
# smoothed_data_combined <- do.call(cbind, smoothed_list)

# # Combine smoothed data into a multivariate functional data object
# # Stack the functional data using fd utilities
# fd_list_combined <- do.call(fd, lapply(smoothed_list, function(x) x$coefs))

# Step 1: Extract coefficients from each smoothed fd object
coef_matrix <- do.call(cbind, lapply(smoothed_list, function(fd_obj) fd_obj$coefs))

# Step 2: Use the same basis as before to create a new multivariate functional data object
combined_fd <- fd(coef_matrix, basis)

# Step 3: Perform functional PCA
fpca_result <- pca.fd(combined_fd, nharm = 4)  # Extract first 4 principal components

# Step 4: Visualize the principal components
plot(fpca_result$harmonics, main = "Multivariate Functional PCA (Study 2)")

# Step 5: Display variance explained by each principal component
fpca_result$varprop
```

```{r}
scores <- fpca_result$scores

plot(scores[,1], scores[,2], 
     xlab = "PC1 Scores", ylab = "PC2 Scores",
     main = "Participant Scores on PC1 vs PC2",
     pch = 19, col = "darkorange")
abline(h = 0, v = 0, lty = 2)
```

```{r}
# Compare PC1 loadings across capacities
plot(fpca_result$harmonics[1], col = "blue", lwd = 2, main = "PC1 Across Capacities")
plot(fpca_result$harmonics[2], col = "red", lwd = 2, main = "PC2 Across Capacities")
plot(fpca_result$harmonics[3], col = "green", lwd = 2, main = "PC3 Across Capacities")
plot(fpca_result$harmonics[4], col = "purple", lwd = 2, main = "PC4 Across Capacities")
```

```{r}
# Merge participant scores with the original data
participant_scores <- data.frame(
  ResponseId = unique(d2_cap_ratings$ResponseId),
  PC1_Score = fpca_result$scores[,1],
  PC2_Score = fpca_result$scores[,2],
  PC3_Score = fpca_result$scores[,3],
  PC4_Score = fpca_result$scores[,4])

# Merge with capacity data
merged_data <- d2_cap_ratings %>%
  left_join(participant_scores, by = "ResponseId")
```

```{r, fig.width = 10}
# Plot how PC1 scores relate to specific capacities (example with 'reasoning_about_things')
merged_data %>%
  mutate(score_cat = case_when(
    PC1_Score < 0 ~ "low",
    PC1_Score >= 0 ~ "high")) %>%
  mutate(capacity = str_wrap(gsub("_", " ", capacity), 20)) %>%
  ggplot(aes(x = target_year, y = response, color = score_cat)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor)))) +
  geom_smooth() +
  # geom_jitter(height = 0, width = 0.2, alpha = 0.05, size = 0.1) +
  # scale_color_gradient(low = "red", high = "blue") +
  theme_minimal() +
  labs(title = "PC1 Scores vs. Reasoning Ratings Across Ages",
       color = "PC1 score (cutoff: 0)",
       x = "Target Age (years)", y = "Rating")
```

```{r, fig.width = 10}
# Plot how PC2 scores relate to specific capacities (example with 'reasoning_about_things')
merged_data %>%
  mutate(score_cat = case_when(
    PC2_Score < 0 ~ "low",
    PC2_Score >= 0 ~ "high")) %>% 
  mutate(capacity = str_wrap(gsub("_", " ", capacity), 20)) %>%
  ggplot(aes(x = target_year, y = response, color = score_cat)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor)))) +
  geom_smooth() +
  # geom_jitter(height = 0, width = 0.2, alpha = 0.05, size = 0.1) +
  # scale_color_gradient(low = "red", high = "blue") +
  theme_minimal() +
  labs(title = "PC2 Scores vs. Reasoning Ratings Across Ages",
       color = "PC2 score (cutoff: 0)",
       x = "Target Age (years)", y = "Rating")
```

```{r, fig.width = 10}
# Plot how PC3 scores relate to specific capacities (example with 'reasoning_about_things')
merged_data %>%
  mutate(score_cat = case_when(
    PC3_Score < 0 ~ "low",
    PC3_Score >= 0 ~ "high")) %>% 
  mutate(capacity = str_wrap(gsub("_", " ", capacity), 20)) %>%
  ggplot(aes(x = target_year, y = response, color = score_cat)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor)))) +
  geom_smooth() +
  # geom_jitter(height = 0, width = 0.2, alpha = 0.05, size = 0.1) +
  # scale_color_gradient(low = "red", high = "blue") +
  theme_minimal() +
  labs(title = "PC3 Scores vs. Reasoning Ratings Across Ages",
       color = "PC3 score (cutoff: 0)",
       x = "Target Age (years)", y = "Rating")
```

```{r, fig.width = 10}
# Plot how PC4 scores relate to specific capacities (example with 'reasoning_about_things')
merged_data %>%
  mutate(score_cat = case_when(
    PC4_Score < 0 ~ "low",
    PC4_Score >= 0 ~ "high")) %>% 
  mutate(capacity = str_wrap(gsub("_", " ", capacity), 20)) %>%
  ggplot(aes(x = target_year, y = response, color = score_cat)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor)))) +
  geom_smooth() +
  # geom_jitter(height = 0, width = 0.2, alpha = 0.05, size = 0.1) +
  # scale_color_gradient(low = "red", high = "blue") +
  theme_minimal() +
  labs(title = "PC4 Scores vs. Reasoning Ratings Across Ages",
       color = "PC4 score (cutoff: 0)",
       x = "Target Age (years)", y = "Rating")
```

```{r}
library(fda)  # Ensure functional data analysis tools are available
# Step 1: Reshape the data to wide format (each Capacity × Target Age as a column)
d2_ratings_matrix <- d2_cap_ratings %>%
  dplyr::select(-c(target, factor)) %>%
  unite("capacity_age", capacity, target_year, sep = "_") %>%
  pivot_wider(names_from = capacity_age, values_from = response) %>%
  column_to_rownames("ResponseId") %>%
  as.matrix()

# Extract observed eigenvalues from fPCA result
observed_eigenvalues <- fpca_result$values

# Set up parallel analysis
num_simulations <- 100  # Increase for better stability
random_eigenvalues <- matrix(NA, nrow = length(observed_eigenvalues), ncol = num_simulations)

# Generate randomized functional datasets and compute fPCA on them
set.seed(123)  # For reproducibility
for (i in 1:num_simulations) {
  shuffled_data <- d2_ratings_matrix[sample(nrow(d2_ratings_matrix)), ]  # Shuffle rows
  smoothed_random <- smooth.basis(argvals = ages, y = t(shuffled_data), fdParobj = basis)$fd
  fpca_random <- pca.fd(smoothed_random, nharm = length(observed_eigenvalues))
  random_eigenvalues[, i] <- fpca_random$values
}

# Compute 95th percentile (upper bound) of randomized eigenvalues
random_threshold <- apply(random_eigenvalues, 1, quantile, probs = 0.95, na.rm = TRUE)

# Plot observed vs. randomized eigenvalues
plot(observed_eigenvalues, type = "b", pch = 19, col = "blue",
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Parallel Analysis for fPCA")
lines(random_threshold, type = "b", pch = 19, col = "red", lty = 2)

legend("topright", legend = c("Observed", "Randomized 95% CI"),
       col = c("blue", "red"), lty = c(1, 2), pch = 19)

# Determine optimal number of PCs (where observed > random)
optimal_pcs <- sum(observed_eigenvalues > random_threshold)
cat("Optimal Number of Functional PCs:", optimal_pcs, "\n")

```