### 2025-02-27 supplement/cover.tex
de-anonymize

\begin{titlepage}
    \begin{center}
        \vspace*{1in}
        {\Large Beliefs about the development of mental life: Supplemental Materials} \\[2cm]

        Kara Weisman\\
        \textit{Department of Psychology, University of California, Riverside}\par
        \vspace{0.5cm}
        Lucy S. King\\
        \textit{Department of Psychology and Human Development, Vanderbilt University}\par
        \vspace{0.5cm}
        Kathryn Humphreys\\
        \textit{Department of Psychology and Human Development, Vanderbilt University}\par

    \end{center}
    \vfill
    \textbf{Author Note} \\
    \setlength{\parindent}{2em}
    \indent Kara Weisman: https://orcid.org/0000-0002-3706-7554. Lucy S. King: https://orcid.org/0000-0002-2552-0614. Kathryn Humphreys: https://orcid.org/0000-0002-5715-6597.

    \indent Weisman, King, and Humphreys were affiliated with the Department of Psychology at Stanford University at the time of data collection. Weisman was affiliated with the Department of Psychology at the University of California, Riverside, at the time of submission, and is now affiliated with the Department of Psychology at Harvard University. King and Humphreys are now affiliated with the Department of Psychology and Human Development at Vanderbilt University.

    \indent Study materials, preregistrations, data, analysis scripts, and links to the fully reproducible manuscript and supplemental materials are openly available via the project’s Open Science Framework page (https://osf.io/xrznd/; DOI: 10.17605/OSF.IO/XRZND). Funding was provided by the William R. and Sara Hart Kimball Stanford Graduate Fellowship (Weisman); the National Science Foundation (Weisman and King: Graduate Research Fellowships; Humphreys: CAREER 2042285), the National Institutes of Health (King: F32 HD105385), and the Jacobs Foundation (Humphreys: Early Career Research Fellowship 2017-1261-05).

    \indent Correspondence should be addressed to Kara Weisman at kgweisman@gmail.com.
\end{titlepage}